### 2022-04-06 outputs/paper/paper.Rmd
add abstract

---
title: "Agriculture and combustion, are two major sources of particular matter(PM2.5) that are associated with premature deaths in Canada"
author: 
  - Zhongyu Huang
thanks: "Code and data are available at: https://github.com/zhongyuhuang/Household-air-pollution-attributable-deaths-Analysis.git"
date: "April 6, 2022"
abstract: "The silent killer, PM2.5, was ranked the fifth risk factor for cardiovascular disease and lung cancer. Approximately 4.5 million individuals died due to exposure to PM2.5 globally every year. This report aims to model and predict the trend of premature deaths resulting from PM2.5 in Canada by using the linear regression model. The results indicate that the number of premature deaths can be explained by man-made emissions from agriculture and combustion. Increasing emissions from these two sources are significantly related to increases in deaths."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(modelsummary)
library(tidyverse)
library(ggplot2)
library(car)
library(dplyr)
f <- file.choose()
APD <- read.csv(f)
```

# Introduction

 The government of Canada reported that an estimated 15,300 premature deaths are linked to air pollution every year. That is, approximately 5% of the total number of deaths reported in 2020. The health effects of air pollution have been subject to intense study in recent years. Exposure to pollutants such as airborne participate matter and ozone has been associated with increases in mortality and hospital admissions due to respiratory and cardiovascular disease. These effects have been found in short-term studies, which relate day-to-day variations in air pollution and health, and long-term studies, which have followed cohorts of exposed individuals over time.

# Data

Our data is of Air pollution (Figure \@ref(fig:pd)).

```{r pd, fig.cap="Premature deaths caused by fine particulate matter", echo = FALSE}
#EDA

for(i in 9:18){
hist(APD[,i], main=paste0("Histogram of ", names(APD)[i]), xlab=names(APD)[i], cex.main=0.75)
}
```
```{r,include=FALSE, echo=FALSE}

APD2 <- APD %>% mutate (Premature.deaths=Premature.deaths+0.5, Waste=Waste+0.5, Agriculture=Agriculture+0.5,Road.Transport=Road.Transport+0.5, Miscellaneous=Miscellaneous+0.5,Power.stations=Power.stations+0.5, Other.Mobile.Sources=Other.Mobile.Sources+0.5, Combustion=Combustion+0.5, Industrial.processes.and.product.use=Industrial.processes.and.product.use+0.5)

APD3<- APD2 %>% select(`Premature.deaths`, Waste, Agriculture, `Road.Transport`, Miscellaneous, `Power.stations` , `Other.Mobile.Sources` , Combustion , `Industrial.processes.and.product.use`)

APD4 <- APD3 %>% mutate(Premature.deaths = log(Premature.deaths), Power.stations=log(Power.stations))
Model2 <- lm(Premature.deaths ~ Waste + Agriculture + Road.Transport + Miscellaneous + Power.stations + Other.Mobile.Sources + Combustion + Industrial.processes.and.product.use, data=APD4)

Model1 <- lm(Premature.deaths ~ Waste + Agriculture + Road.Transport + Miscellaneous + Power.stations + Other.Mobile.Sources + Combustion + Industrial.processes.and.product.use, data=APD3)

Model3 <- lm(Premature.deaths ~ Waste + Agriculture + Road.Transport + Miscellaneous + Power.stations  + Combustion + Industrial.processes.and.product.use, data=APD4)
#vif(Model3)
#summary(Model3)$adj.r.squared
#summary(Model2)$adj.r.squared
Model4 <- lm(Premature.deaths ~ Waste + Agriculture  + Miscellaneous + Power.stations  + Combustion + Industrial.processes.and.product.use, data=APD4)

Model6 <- lm(Premature.deaths ~ Waste + Agriculture  + Miscellaneous  + Combustion+  Road.Transport, data=APD4)

Model7 <- lm(Premature.deaths ~  Agriculture  + Miscellaneous + Combustion, data=APD4)

#check condition1 and condition2
pairs(APD[, -c(1:9,18)])
plot(APD$Premature.deaths ~ fitted(Model1), main="Y versus Yhat", ylab="response", xlab="yhat")
abline(a = 0, b = 1)
lines(APD$Premature.deaths ~ fitted(Model1), lty=2)
#Transform

boxCox(Model1)
p <- powerTransform(cbind(APD3[,-c(1,2)]))
summary(p)

plot(APD4$Premature.deaths ~ fitted(Model2))
plot(Model2)
plot(Model1)

mod1 <- lm(Premature.deaths ~ Waste + Agriculture +Road.Transport+Miscellaneous +Power.stations, data=APD4)
anova(mod1, Model2)
plot(resid(mod1) ~ fitted(mod1))
plot(resid(mod1) ~ APD4$Waste)
plot(resid(mod1) ~ APD4$Agriculture)
qqnorm(resid(mod1))
qqline(resid(mod1))

n <- nrow(APD4)
p <- length(coef(Model2))-1
Hcut <- 2*((p+1)/n)
DFFITScut <- 2*sqrt((p+1)/n)
DFBETAcut <- 2/sqrt(n)
Dcut <- qf(0.5, p+1, n-p-1)
# identify the leverage points
h <- hatvalues(Model2)
which(h>Hcut)
h
r <- rstandard(Model2)
which(r < -2 | r > 2)

for(i in 9:18){
boxplot(APD[,i], main=paste0("Boxplot of ", names(APD)[i]), xlab=names(APD)[i]
, horizontal=T)
}
```
```{r,include=FALSE,echo=FALSE}
#install.packages("modelsummary")

p<-modelsummary(list(Model1,
                  Model2, 
                  Model7),
             fmt = 2,
             title = "Explaining marathon time based on five-kilometer run times and weather features")
p
```

Talk more about it.

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}

```

Talk way more about it.

# Model

\begin{equation}
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
\end{equation}

Equation \@ref(eq:bayes) seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional details

\newpage

# References